---
title: "Supplementary Files"
output:
  html_document:
    df_print: paged
    toc: yes
    toc_depth: '3'
    toc_float: true
    code_folding: hide
---

```{=html}
<style>
div.blue { background-color:#e6f0ff; border-radius: 5px; padding: 10px;}
</style>
```
::: {.blue}
-   Goal:

    -   setting up common names across all files
:::

# Data Prep
## Libraries

```{r}
## Install the following packages 
library(stringr)
library(readxl)
library(data.table)
library(dplyr)
library(here)
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# knitr::opts_knit$set(root.dir = "/Users/sakaizarasolofomananarajery/Desktop/Research/Hendriks Folder/Biotransformers/")
```
## Insert Interested Outcomes
```{r}
outcomes <- c("SBP_total","DBP_total","Insulin","Chol_T","TG","LDL","HDL","CRP","ApoB")
foodtype <- "salmon"
datatype <- "aqueous_bio"
```

## Reading in Data
Read in Targeted File as "food"
```{r}


food <-read.table("AQ pos unique to salmon predicted metabolites_found in plasma_Profinder.txt", 
                 sep="\t", header=T)
```

```{r}
# Metabolite Identifiers (Mass, Retention, Compound)
food <- as.data.frame(t(food))
food <- cbind(rownames(food),food)

colnames(food) <- food["Compound",]
food <- food[-1,]
rownames(food) <- NULL
colnames(food)[1] <- "StudyID"

metabolites <- colnames(food)[2:length(food)] # identify metabolite names

# Create a new dataframe with information about metabolites
met_info <- t(food[!grepl("raw", food$StudyID),])
rownames(met_info) <- NULL
colnames(met_info) <- met_info[1,]
met_info <- met_info[-1,]
met_info <- as.data.frame(met_info[,c("Alignment.Value","Annotations", "Compound.Name", "Formula", "Ion.Species",
                                      "Mass","Retention.Time")])
```

Removing Unwanted Rows(Rows Not Containing "raw" in Their Names)
```{r}
food <- food[grepl("raw", food$StudyID),]
#food$StudyID <- gsub("report.results_", "", food[,1], fixed = T)
#food$StudyID <- gsub("(raw)", "", food[,1], fixed = T)
food$StudyID <- gsub("X","",food$StudyID)
food$StudyID <- gsub(".raw.","",food$StudyID)
```


## Import AQ Key Sample
```{r}
#fil_path <- here("data", "original_files","Med_Diet_AQ_Sample_Key.xlsx")
key<- read_excel("./Med_Diet_AQ_Sample_Key.xlsx") 
key$`Plasma Sample ID`[163:182] <- key$`Instrument Sample Name`[163:182]
key$`Instrument Batch` <- gsub("Batch_","",key$`Instrument Batch`)

key <- key[!grepl("Prep_Blank",key$`Plasma Sample ID`),]
#key$`Plasma Sample ID` <- gsub(".*Instrument", "Instrument", key$`Plasma Sample ID`)
key$`Plasma Sample ID` <- gsub("S44-", "", key$`Plasma Sample ID`, fixed = T)
key$`Plasma Sample ID` <- gsub("-", " ", key$`Plasma Sample ID`, fixed = T)
key$`Plasma Sample ID` <- gsub("  P*"," P", key$`Plasma Sample ID`)
key <- key%>%
  rename(StudyID = "Instrument Sample Name", batch ="Instrument Batch", order ="Instrument Run Order")

joined <- right_join(key,food)%>%
  select(-c("StudyID"))%>%
  rename(StudyID = `Plasma Sample ID`)

food<- joined[, c(3,2,1,4:ncol(joined))]
```



## Read Supplementary Files
reading in "info_file", "batch_file", "complete_health_outcome". Info File Contains Following Info: "Period", "Trial", "Time", "Age", and "Sex"
```{r}

Info_file <- read.csv("./S44data_MASTER_41 subjects.csv" ,as.is = T)
Info_file <- Info_file[,c("StudyID", "Period", "Trial", "Time","Age", "Sex", "Diet")] 
Info_file <- Info_file[complete.cases(Info_file),] # Removing rows with incomplete data
```

### Health Outcome
```{r}
complete_health_outcome <- read.csv("./S44data_MASTER_41 subjects.csv")
```

#### Only Reading "outcomes" of Health File
```{r}
complete_health_outcome <- complete_health_outcome[,c("StudyID", "Period", "Trial", outcomes)]

## replacing "." with NA
complete_health_outcome[complete_health_outcome=="."] <- NA
```

## Setting Up a Common ID
Each File has different names and a common names is needed
```{r}
Info_file[1:5,1]
#batch_file[1:5,1]
food[1:3,1]
```

### Info_file
```{r}
Info_file$StudyID <- gsub("S44-","", Info_file$StudyID, fixed = TRUE)
Info_file$StudyID <- paste(Info_file$StudyID, Info_file$Time, Info_file$Trial, sep = " ")
Info_file$StudyID <- gsub("p","P",Info_file$StudyID)
```

### food
```{r}
food$day_numb <- ifelse(str_detect(food$StudyID,"Blank_Plus_Day_"), str_extract(food$StudyID,"[0-9]+"),NA)
```

Checking Results
```{r}
Info_file$StudyID[1:5]
#batch_file$StudyID[1:5]
food$StudyID[1:5]
```

### Merging "food","Info_file","batch_file"
```{r}

new_merged <- merge(Info_file,food, by="StudyID",all=T)

## remove incorrect duplicated QC values
QC_ID <- new_merged[grepl("Instrument",new_merged$StudyID),]
QC_ID <- QC_ID[QC_ID$batch==QC_ID$day_numb,]  
new_merged <- new_merged[is.na(new_merged$day_numb),]
#new_merged <- rbind(new_merged,QC_ID)
new_merged$day_numb <- NULL
```

### Adding two Health Outcomes of "HDL/LDL", and "Chol_T/HDL"
```{r}
complete_health_outcome$HDL_LDL <- complete_health_outcome$HDL/complete_health_outcome$LDL

complete_health_outcome$Chol_T_HDL <- complete_health_outcome$Chol_T/complete_health_outcome$HDL
```

### Creating "health_outcome", which is the Difference in Outcomes within Each Trial
```{r}
outcomes <- c("SBP_total","DBP_total","Insulin","Chol_T","TG","LDL","HDL","CRP","ApoB","HDL_LDL","Chol_T_HDL")

complete_health_outcome[,outcomes] <- apply(complete_health_outcome[,outcomes],2,function(x) as.numeric(as.character(x)))

complete_health_outcome$StudyID <- gsub(".*-","",complete_health_outcome$StudyID)

names_health <- as.character(unique(complete_health_outcome$StudyID))
tmp.health <- diff_health <- tmp.health.dat <- c()

for(j in names_health){
  tmp.health <- complete_health_outcome[complete_health_outcome$StudyID==j,]
  
  for(k in 1:2){
    if(sum(tmp.health$Trial == k) == 2){
      tmp.health.dat <- tmp.health[tmp.health$Trial == k,] 
      
      tmp.health.dat <- (tmp.health.dat[grepl(max(tmp.health.dat$Period),tmp.health.dat$Period),outcomes])-
        (tmp.health.dat[grepl(min(tmp.health.dat$Period),tmp.health.dat$Period),outcomes])

      tmp.health.dat <- cbind(StudyID=j, Trial=k, tmp.health.dat)
      
      diff_health <- rbind(diff_health, tmp.health.dat)
    }
  }
}
```

# Writting Out Results

```{r}

file.name <- paste(foodtype, datatype, sep = "_")

write.table(new_merged,
            file=paste("./merged_file_", file.name,".txt",sep = ""),row.names=F, quote = T)
write.table(diff_health,
            file=paste("./health_outcome_", file.name,".txt",sep = ""),row.names=F, quote = T)
write.table(complete_health_outcome,
            file=paste("./complete_health_outcome_", file.name,".txt",sep = ""),
            row.names = F, quote = T)
write.table(met_info,
            file=paste("./met_info_", file.name,".txt",sep = ""),
            row.names = F, quote = T)

```


